Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-06-05 old.tex
Added Allen's changes.  Submitting this version to be safe.

\begin{abstract} Recently a family of promising face recognition algorithms
based on sparse representation and $\ell_1$-minimization ($\ell_1$-min) have been
developed.  These algorithms have not yet seen commercial application, 
largely due to higher computational cost compared to other traditional
algorithms. This paper studies techniques for leveraging the 
massive parallelism available in GPU and CPU hardware to accelerate
$\ell_1$-min based on augmented Lagrangian method (ALM) solvers. 
Due to the low arithmetic intensity of ALM solvers, implementations on both
architectures are primarily limited by memory bandwidth,  
with large problems running faster on the GPU and small problems running faster
on the CPU.  
On both architectures, the proposed implementations significantly outperform
naive library-based implementations. 
At the problem sizes most important for $\ell_1$-min based
face recognition, both architectures achieve comparable performance.
The source code of the
algorithms will be made available for peer evaluation.  \end{abstract}
\vspace{-0.1in}

@@

Figure \ref{fig:alignment_stage_runtime} shows the average run time of the CPU and GPU implementations to align
a query image against all the subject classes separately. We vary the total number of
subject classes to show how the algorithms scale. The plateaus seen in the GPU curve
result from the GPU hardware scheduling the computation of alignment
problems in concurrent batches, but the overall trend is linear as expected.
The manually threaded CPU implementation slightly outperforms the GPU implementation,
and it beats the \emph{naive}
library threaded implementation by a wide margin.
Compared to previously published results, the new implementation can align the query
image with 250 subject classes in under 10 seconds, while the best previous effort
required 1.7 seconds per subject class (i.e. 253 seconds for 250 subjects).

>>>>

\begin{abstract} Recently a family of promising face recognition algorithms
based on sparse representation and $\ell_1$-minimization ($\ell_1$-min) have been
developed.  These algorithms have not yet seen commercial application, 
largely due to higher computational cost compared to other traditional
algorithms. This paper studies techniques for leveraging the 
massive parallelism available in GPU and CPU hardware to accelerate
$\ell_1$-min based on augmented Lagrangian method (ALM) solvers. 
For very large problems the GPU is faster to higher memory bandwidth, while 
for problems that fit in the larger CPU L3 cache, the CPU is faster.
At the problem sizes most important for $\ell_1$-min based
face recognition, both architectures achieve comparable performance.
On both architectures, the proposed implementations significantly outperform
naive library-based implementations, as well as published results. 
The source code of the
algorithms will be made available for peer evaluation.  \end{abstract}
\vspace{-0.1in}

@@

Figure \ref{fig:alignment_stage_runtime} shows the average run time of the CPU and GPU implementations to align
a query image against all the subject classes separately. We vary the total number of
subject classes to show how the algorithms scale. The plateaus seen in the GPU curve
result from the GPU hardware scheduling the computation of alignment
problems in concurrent batches, but the overall trend is linear as expected.
The manually threaded CPU implementation slightly outperforms the GPU implementation,
and it beats the \emph{naive}
library threaded implementation by a wide margin.
The new implementation can align the query
image in $\approx 40$ ms per subject, 
while the fastest previously published result \cite{WagnerA2011-PAMI}
required $\approx 600$ ms seconds per subject in a similar setting.
